---
title: "pass_test Data processing 40%-liyu, 20%-collaboration,model prediction 40%-zhaohongqiang "
author: "zhao hongqiang"
date: "6/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(zoo)
library(fpp2)
library(xlsx)
library(tseries)
library(dplyr)
library(tidyr)
proc_data <- function(xlsx,foodnames,indexs) {
datas = read.xlsx(xlsx,sheetIndex = indexs,startRow = 2)
alldata = data.frame(
  vol  = datas[,c(paste("VOL_FRESH." , foodnames,sep = ""))],
  pr = datas[,c(paste("PR_FRESH.",foodnames,sep = ""))],
  date = as.Date(time(ts(datas$PR_FRESH.DOUGHNUTS,frequency = 12 ,start = c(2004,7))))
)%>%
  separate(date ,into=c("year","month","day"),sep = "-",remove = F)
alldata$turnover = alldata$vol*alldata$pr
alldata$quarters = quarters(alldata$date)
return(alldata)
}
proc_ts_data <- function(datas) {
  month_vol = ts(datas[,"vol"],frequency = 12,start = c(2004,7))
  month_turn=ts(datas[,"turnover"],frequency = 12,start = c(2004,7))
  month_pr = ts(datas[,"pr"],frequency = 12,start = c(2004,7))
  
  
  quarters_turn = ts((datas%>% group_by(year,quarters) %>%  
                             summarise(turnover=sum(turnover),.groups = 'drop'))[,'turnover'],frequency = 4,start = c(2004,7))
  quarters_vol = ts((datas%>% group_by(year,quarters) %>%  
                            summarise(vol=sum(vol),.groups = 'drop'))[,'vol'],frequency = 4,start = c(2004,7))
  quarters_pr = ts((datas%>% group_by(year,quarters) %>%  
                           summarise(pr=sum(pr),.groups = 'drop'))[,'pr'],frequency = 4,start = c(2004,7))
  
  years_turn = ts(
    (datas%>% group_by(year) %>%  
       summarise(turnover=sum(turnover)))[,'year'],frequency = 1,start = c(2004))
  years_pr = ts(
    (datas%>% group_by(year) %>%  
       summarise(pr=sum(pr)))[,'pr'],frequency = 1,start = c(2004))
  years_vol = ts(
    (datas%>% group_by(year) %>%  
       summarise(vol=sum(vol)))[,'vol'],frequency = 1,start = c(2004))
  
  out = list(
    month_vol,month_pr,month_turn,
    years_turn,years_vol,years_pr,
    quarters_pr,quarters_turn,quarters_vol
  )
  
  return(out)
}
```

# Sales Data Forecast


## 1.Data process

### 1.1 Statistics data at year season and month

```{r}
mass_data = proc_data(xlsx = './Data_for_test.xlsx',"DOUGHNUTS",1)

### food data porcess 
food_data = proc_data(xlsx = './Data_for_test.xlsx',"DOUGHNUTS",2)
### drug data porcess 

drug_data = proc_data(xlsx = './Data_for_test.xlsx',"DOUGHNUTS",3)

```

### 1.2 used TS func convert time series

```{r}
#    1：2：3=month_vol,month_pr,month_turn,
#    4；5；6=years_vol,years_pr,years_turn,
#   7：8：9=quarters_vol,quarters_pr,quarters_turn
mass=proc_ts_data(mass_data)
food= proc_ts_data(food_data)
drug= proc_ts_data(drug_data)
head(food[[2]])
```


## 2Forecast

### 2.1 time regression for year turnover data

```{r}

autoplot(mass[[4]],main = "mass_vol")
autoplot(mass[[5]],main = "mass_pr")
autoplot(mass[[6]],main = "mass_turn")
autoplot(food[[4]],main = "food_vol")
autoplot(food[[5]],main = "food_pr")
autoplot(food[[6]],main = "food_turn")
autoplot(drug[[4]],main = "drug_vol")
autoplot(drug[[5]],main = "drug_pr")
autoplot(drug[[6]],main = "drug_turn")
## used tslm for year data
fc_tsml <- function(datas) {
  fc_year <- tslm(datas ~ trend )
fcast <- forecast(datas, h=3)
autoplot(fcast)
 return(fcast)
}
fc_year_mass_vol = fc_tsml(mass[[4]])
autoplot(fc_year_mass_vol,main="MASS Year VOL Forecast")
fc_year_mass_pr = fc_tsml(mass[[5]])
autoplot(fc_year_mass_pr,main="MASS Year PR Forecast")
fc_year_mass_turn = fc_tsml(mass[[6]])
autoplot(fc_year_mass_turn,main="MASS Year TURNOVER Forecast")

fc_year_food_vol = fc_tsml(food[[4]])
autoplot(fc_year_food_vol,main="FOOD Year VOL Forecast")
fc_year_food_pr = fc_tsml(food[[5]])
autoplot(fc_year_food_pr,main="FOOD Year PR Forecast")
fc_year_food_turn = fc_tsml(food[[6]])
autoplot(fc_year_food_turn,main="FOOD Year TURNOVER Forecast")

fc_year_drug_vol = fc_tsml(drug[[4]])
autoplot(fc_year_drug_vol,main="DRUG Year VOL Forecast")
fc_year_drug_pr = fc_tsml(drug[[5]])
autoplot(fc_year_drug_pr,main="DRUG Year PR Forecast")
fc_year_drug_turn = fc_tsml(drug[[6]])
autoplot(fc_year_drug_turn,main="DRUG Year TURNOVER Forecast")

```
### 2.2 Forecast season Data at ARIMA model

#### Manual ARIMA analysis of drug turnover data to predict outcomes over three years

```{r}
autoplot(drug[[9]])
ggseasonplot(drug[[9]])
ndiffs(drug[[9]])
nsdiffs(drug[[9]])
# the result shows need 1 seasonal diff and 1 diff
drug[[9]]%>% diff(lag=4) %>% diff()%>% ggtsdisplay(main="")
auto.arima(drug[[9]])
# the auto arima used 1,0,0,  seasonal 0,1,0
drug[[9]]%>%
  Arima(order=c(1,1,0),seasonal=c(0,1,0))
### 1,1,0 AICc = 393.55
drug[[9]]%>%
  Arima(order=c(0,1,1),seasonal=c(0,1,0))
### 0,1,1   AICc=393.53 

drug[[9]]%>%
  Arima(order=c(0,1,1),seasonal=c(0,1,0)) %>%
  residuals() %>%
  ggtsdisplay()
# the 0,1,0 AICc value is smaller so choose 0,1,0
drug[[9]]%>%
  Arima(order=c(0,1,1),seasonal=c(0,1,0))->fit3
checkresiduals(fit3)
fit3 %>% forecast(h=12) %>% autoplot()
```
 
###  used auto model Forecast data 

```{r}
autoplot(mass[[7]])
autoplot(mass[[8]])
autoplot(mass[[9]])
autoplot(food[[7]])
autoplot(food[[8]])
autoplot(food[[9]])
autoplot(drug[[7]])
autoplot(drug[[8]])
autoplot(drug[[9]])

arima_model <- function(datas) {
  arima_data =  auto.arima(datas)
  fc_arima = forecast(
  arima_data,
  h = 12
)
  return(fc_arima)
}
fc_season_mass_vol = arima_model(mass[[7]])
autoplot(fc_season_mass_vol,main="MASS Season VOL Forecast at ARIMA model")
fc_season_mass_pr = arima_model(mass[[8]])
autoplot(fc_season_mass_pr,main="MASS Season PR Forecast at ARIMA model")
fc_season_mass_turn = arima_model(mass[[9]])
autoplot(fc_season_mass_turn,main="MASS Season TURNOVER Forecast at ARIMA model")
fc_season_food_vol = arima_model(food[[7]])
autoplot(fc_season_food_vol,main="FOOD Season VOL Forecast at ARIMA model")
fc_season_food_pr = arima_model(food[[8]])
autoplot(fc_season_food_pr,main="FOOD Season PR Forecast at ARIMA model")
fc_season_food_turn = arima_model(food[[9]])
autoplot(fc_season_food_turn,main="FOOD Season TURNOVER Forecast at ARIMA model")
fc_season_drug_vol = arima_model(drug[[7]])
autoplot(fc_season_drug_vol,main="DRUG Season VOL Forecast at ARIMA model")
fc_season_drug_pr = arima_model(drug[[8]])
autoplot(fc_season_drug_pr,main="DRUG Season PR Forecast at ARIMA model")
fc_season_drug_turn = arima_model(drug[[9]])
autoplot(fc_season_drug_turn,main="DRUG Season TURNOVER Forecast at ARIMA model")

```





### 2.3 Forecast month Data uesd Holt-Winters multiplicative model

#### Divide the data into a test set training set and use the aucc function to check the accuracy of the results

```{r}
turn_train <- subset(drug[[1]],
                         end = length(drug[[1]]) - 36)
turn_test <- subset(drug[[1]],
                        start = length(drug[[1]]) - 35)
hw_mul_turn_train <- hw(turn_train,
                            h = 36,
                            seasonal = "multiplicative")
autoplot(hw_mul_turn_train)
accuracy(hw_mul_turn_train, turn_test)
checkresiduals(hw_mul_turn_train)



turn_train <- subset(drug[[2]],
                         end = length(drug[[2]]) - 36)
turn_test <- subset(drug[[2]],
                        start = length(drug[[3]]) - 35)
hw_mul_turn_train <- hw(turn_train,
                            h = 36,
                            seasonal = "multiplicative")
autoplot(hw_mul_turn_train)
accuracy(hw_mul_turn_train, turn_test)
checkresiduals(hw_mul_turn_train)


turn_train <- subset(drug[[3]],
                         end = length(drug[[3]]) - 36)
turn_test <- subset(drug[[3]],
                        start = length(drug[[3]]) - 35)
hw_mul_turn_train <- hw(turn_train,
                            h = 36,
                            seasonal = "multiplicative")
autoplot(hw_mul_turn_train)
accuracy(hw_mul_turn_train, turn_test)
checkresiduals(hw_mul_turn_train)


############
turn_train <- subset(mass[[1]],
                         end = length(mass[[1]]) - 36)
turn_test <- subset(mass[[1]],
                        start = length(mass[[1]]) - 35)
hw_mul_turn_train <- hw(turn_train,
                            h = 36,
                            seasonal = "multiplicative")
autoplot(hw_mul_turn_train)
accuracy(hw_mul_turn_train, turn_test)
checkresiduals(hw_mul_turn_train)



turn_train <- subset(mass[[2]],
                         end = length(mass[[2]]) - 36)
turn_test <- subset(mass[[2]],
                        start = length(mass[[3]]) - 35)
hw_mul_turn_train <- hw(turn_train,
                            h = 36,
                            seasonal = "multiplicative")
autoplot(hw_mul_turn_train)
accuracy(hw_mul_turn_train, turn_test)
checkresiduals(hw_mul_turn_train)


turn_train <- subset(mass[[3]],
                         end = length(mass[[3]]) - 36)
turn_test <- subset(mass[[3]],
                        start = length(mass[[3]]) - 35)
hw_mul_turn_train <- hw(turn_train,
                            h = 36,
                            seasonal = "multiplicative")
autoplot(hw_mul_turn_train)
accuracy(hw_mul_turn_train, turn_test)
checkresiduals(hw_mul_turn_train)

######
turn_train <- subset(food[[1]],
                         end = length(food[[1]]) - 36)
turn_test <- subset(food[[1]],
                        start = length(food[[1]]) - 35)
hw_mul_turn_train <- hw(turn_train,
                            h = 36,
                            seasonal = "multiplicative")
autoplot(hw_mul_turn_train)
accuracy(hw_mul_turn_train, turn_test)
checkresiduals(hw_mul_turn_train)



turn_train <- subset(food[[2]],
                         end = length(food[[2]]) - 36)
turn_test <- subset(food[[2]],
                        start = length(food[[3]]) - 35)
hw_mul_turn_train <- hw(turn_train,
                            h = 36,
                            seasonal = "multiplicative")
autoplot(hw_mul_turn_train)
accuracy(hw_mul_turn_train, turn_test)
checkresiduals(hw_mul_turn_train)


turn_train <- subset(food[[3]],
                         end = length(food[[3]]) - 36)
turn_test <- subset(food[[3]],
                        start = length(food[[3]]) - 35)
hw_mul_turn_train <- hw(turn_train,
                            h = 36,
                            seasonal = "multiplicative")
autoplot(hw_mul_turn_train)
accuracy(hw_mul_turn_train, turn_test)
checkresiduals(hw_mul_turn_train)

```



#### Automated forecasting using data models directly


```{r}
autoplot(mass[[1]],main = "mass_vol")
autoplot(mass[[2]],main = "mass_pr")
autoplot(mass[[3]],main = "mass_turn")
autoplot(food[[1]],main = "food_vol")
autoplot(food[[2]],main = "food_pr")
autoplot(food[[3]],main = "food_turn")
autoplot(drug[[1]],main = "drug_vol")
autoplot(drug[[2]],main = "drug_pr")
autoplot(drug[[3]],main = "drug_turn")

hw_model <- function(datas) {
  hw_mul <- hw(datas,
                            h = 36,
                            seasonal = "multiplicative")
  return(hw_mul)
}
fc_month_mass_vol = hw_model(mass[[1]])
autoplot(fc_month_mass_vol,main="MASS month VOL Forecast at Holt-Winters multiplicative model")
fc_month_mass_pr = hw_model(mass[[2]])
autoplot(fc_month_mass_pr,main="MASS month PR Forecast at Holt-Winters multiplicative model")
fc_month_mass_turn = hw_model(mass[[3]])
autoplot(fc_month_mass_turn,main="MASS month TURNOVER Forecast at Holt-Winters multiplicative model")
fc_month_food_vol = hw_model(food[[1]])
autoplot(fc_month_food_vol,main="FOOD month VOL Forecast at Holt-Winters multiplicative model")
fc_month_food_pr = hw_model(food[[2]])
autoplot(fc_month_food_pr,main="FOOD month PR Forecast at Holt-Winters multiplicative model")
fc_month_food_turn = hw_model(food[[3]])
autoplot(fc_month_food_turn,main="FOOD month TURNOVER Forecast at Holt-Winters multiplicative model")
fc_month_drug_vol = hw_model(drug[[1]])
autoplot(fc_month_drug_vol,main="DRUG month VOL Forecast at Holt-Winters multiplicative model")
fc_month_drug_pr = hw_model(drug[[2]])
autoplot(fc_month_drug_pr,main="DRUG month PR Forecast at Holt-Winters multiplicative model")
fc_month_drug_turn = hw_model(drug[[3]])
autoplot(fc_month_drug_turn,main="DRUG month TURNOVER Forecast at Holt-Winters multiplicative model")
```










